\documentclass[titlepage]{article}
\usepackage[utf8]{inputenc}
\renewcommand{\abstractname}{Declaration}

\title{\textbf{Battleship} \\ CS251 : Software Systems Lab}
\author{Rahul Bhardwaj, Anuj Diwan, Soumya Chatterjee \\ 170050012, 170070005, 170070010}
\date{November 2018}

\begin{document}

\maketitle

\begin{abstract}
I acknowledge and understand that plagiarism is wrong. This project is my own work, or my group’s own unique group project. I acknowledge that copying someone else's work, or part of it, is wrong, and that submitting identical work to others constitutes a form of plagiarism.

\tableofcontents

\break

\section{Introduction and Motivation}
\subsection{Introduction}
% About the project %
We have implemented an online version of the classic multiplayer game Battleship with a twist which can be played on a web browser. The game supports multiple simultaneous game sessions and has unconventional features like T and L shaped ships and timed attacks. Opponents are also provided with an interface through which they can chat while playing the game.

\subsection{Motivation}
Having worked with various different languages, frameworks and tools throughout our course `Software Systems Lab', we were eager to apply all that we had learnt to create something substantial and complete. The course project gave us the valuable opportunity to explore new frameworks while applying the knowledge and experience gained throughout the course.

Our main motivation behind taking up this project was to to get an experience of working on a Client Server architecture, WebSockets, learning database management. We also wanted to explore different communication protocols.
\section{Features}

\section{Implementation Details}
\subsection{Software and Tools Used}
We used the Django automatic server for automatic handling of database operations, redis server for using django channels to establish persistent connections. SQlite was used for handling the database. 
\subsection{Login and Logout Features}
The inbuilt django auth app was used to implement login, logout and and signup feature. The default form was used for login and logout and custom form for signup. The custom signup form was built by extending the existing user creation from, with additional corresponding to the fields of the profile object. We used django-crispy-forms to bootsrap the forms. 
\section{Backend}
 The database handling was done using Sqlite. We created a "profile" class to store user objects which contains information like username, bio, location, date of birth and a boolean which store whether the player is currently available or not. We have defined 2 views in the pairing app- profile view(is exactly what is sounds like:) and the "listAvailable" to see the list of players currently available. Upon logging the user will be directed the profile view, where he can click on the link which will take him to the listAvailable view. The list of players which are currently active will get updated automatically if a user logs out or begins playing. 
\subsection{Websocket Connection}
The list of players which are currently active(i.e ready to play) needed to be updated automatically. To implement this, we established a web-socket connection by routing it to the Redis server and a persistent connection was formed along with associated consumer object. This was implemented using django channels. We wrote a consumer called ChatConsumer to handle the events of players sending, accepting and rejecting requests. When a user logins in, the information that the player now is available is sent to the entire group of channels and its "isAvaliable" attribute is set to true. The same thing happens when a user logs out. For other types of messages, the type of message that has been sent is stored in the purpose attribute of the message. 
\subsubsection{Recieving Messages}
If the message contains the "loggedIn attribute" then it corresponds to a user logging in/out, then there will be no purpose attribute and the only action which will be performed is the list of users getting updated. If the message contains the logged\_in attribute, we show that that user also in the list, else we hide it.  On the other hand if the message does not contain the "loggedIn" attribute, then we check the purpose attribute of the message. If the purpose attribute is "Send Request", then we show the model displaying that a play request has been sent. If the purpose attribute is "Accept\_Request", then we display the message that his request was accepted by the other player and then direct them to the play pages. If the purpose attribute was "Reject\_Request", then we display a message saying that his request was rejected.  His decision will also be conveyed to the sender of the message. We handled all this using the javascript function socket.onmessage in list.html

\subsubsection{Sending Messages}
In the list.html we defined a function to handle the events of a user clicking on another user and thereby sending a request. The message was again sent as a JSON containing "from","to","purpose" and "type" attributes. The javascript function sends this message over WebSocket to the ChatConsumer. The Chat Consumer upon receiving the message forwards it to the entire group of channels. The ChatConsumer's of the receiving channel then sends the message over websocket to the javascript, which is then handled in the way described in the above paragraph.

\section{Frontend}
The front-end was implemented using the Javascript library React. 
\subsection{Placing the Ships}
In the placing.html file we have created a class called board which will store the current state of the board and defined a function to handle the onclick event. It contains a variable to store which all points on the borad and another variable which stores the points which a particular ship covers. The render function is designed to display the board and give the user options to select ships and place them on the board. We have also created functions which handle events related to placing of ships. For all the ships, we have decided on keeping a point as pivot and then allowing rotations about that point. There are 4 possible orientations for every ship(though some of them may be same for a few pieces). Clicking on a ship selects it and clicking on it subsequent times, rotates it. The player can choose a ship, its orientation and the point where to keep its pivot, after which the function "handleClick" will be called. Based on the ship and its orientation we push all the points covered by that particular ship into an array and then check if any of them goes outside the board and if all the ships have been placed. If both these conditions are met, we send a message through the websocket to django in order to update the database. We also display the number of ships of each type which have been successfully placed.

\subsection{The Actual Game}
To enable communication between the players, we have again made use of websocket connections. As soon as the players get redirected to the play page, the socket connection gets established. Consumers.py contains a "SubmitConsumer" class, whose member functions handle various events related to the game. If a player disconnects in while the game is going on, submit consumer sends a message to the group that the player has disconnected. The message has a purpose attribute which is set to "left\_game" for a message of this kind. The file placing.html contains a function to handle the message recieved from the websocket. The function checks the purpose attribute of the message. If it is set to "left\_game", then it displays a message saying that the opponent has left the game.
\subsection{Shooting}
The "game.html" file contains a Board class. It renders the board and also the number of ships which are left to be sunk by the current player. It has an HandleClick function, which when called updates all the info and sends a message to django in order to update the state. Django on receiving the message updates the corresponding game object and the sends the message to javascript and also back to the group, indicating that the data has been updated. 
For handling the message at the receiver's end, we have socket.onmessage function, which gets called when a message is received. It internally then calls updateState, which updates information like the state of the board and the current player after checking whether the player receiving the message is player 1 or player 2.
\subsection{Allowing Chat between players}
The idea is again almost the same. "Consumers.py" contains a class named "GameChatConsumer" to handle chat messages.Connect and disconnect functions handle joining and leaving of players from the chat room. The reciev function on recieving from the websocket echoes the message to the entire group. The chat message function on recieving a message check if the purpose attribute is "Game\_Chat" and if the game\_is equal to self.game\_id, in which case, it sends the message to the javascript. To handle the javascript part we have defined a class called GameChat. The class contains 2 string array's in it state. One of them store the messages and the other stores the sender. It componentDidMOunt functions handles the receiving of messages. When a message is recieved through the websocket, it pushes the message and sender into the respective arrays. The render function of the class displays the message the along with the senders. 
\section{Contributions}

\end{document}
